---
title: "CellPrecursors_analysis"
author: "Joana P. Bernardes"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(devtools)
library(SingleR)
library(knitr)
library(RColorBrewer)
library(org.Hs.eg.db)
library(topGO)
library(plyr)
library(lmerTest)
```

##Load object extracted from merged object
```{r}
Precursors <- readRDS('CellPrecursors.rds')
```


## Seurat analysis
# Normalize object
```{r}
Precursors <- NormalizeData(Precursors)
```

# Scale object using all genes
```{r}
all.genes<-rownames(Precursors)
Precursors <- ScaleData(Precursors, features = all.genes)
```

## Find object variable genes 
```{r}
Precursors<-FindVariableFeatures(Precursors, selection.method = "vst", nfeatures = 2000)
```

# Run PCA
# Use elbow plot to define number of dimensions
```{r}
Precursors <- RunPCA(Precursors,npcs = 80, ndims.print = 1:5)
ElbowPlot(Precursors, ndims = 80)
```

# run UMAP
```{r}
Precursors <- RunUMAP(object = Precursors, dims = 1:80)
```



# plot basic UMAPs
```{r}
p1<-DimPlot(object = Precursors, reduction = "umap",pt.size = 2,label = TRUE, group.by = "orig.ident")

p2<-DimPlot(object = Precursors, reduction = "umap",pt.size = 2,label = TRUE, group.by = "Patient") 

color<-brewer.pal(n = 11, name = "BrBG")
p3<-DimPlot(object = Precursors, reduction = 'umap', label = FALSE, group.by ="Timepoint", pt.size = 2, 
            order = c('TA', 'TA2', 'TB', 'TC', 'TE', 'rec'),
            cols=rev(color))

p4<-DimPlot(object = Precursors, reduction = "umap",pt.size = 2, label = F, group.by = "Pseudotime",
            order=c( "0",'7',"1",  "3", "4","5","6", '2'),
            cols = c("#7031B9", "#FDC077", "#F99B1C", "#E65826",'#A04E9E',"#E84F8C",'#51BBFE','#A7A9AC'))
CombinePlots(plots = list(p1,p2,p3,p4), ncol=1)
```


## Clustering
# Find neighbours
```{r}
Precursors <- FindNeighbors(Precursors,  dims = 1:80)
```

# Calculate clusters
```{r}
Precursors <- FindClusters(Precursors, resolution = 0.1)
```



# Identify signature genes per Pseudotime
```{r}
Idents(Precursors)<-Precursors$Pseudotime
my_levels <- c("0","1",'2', "3", "4","5","6", '7')
Idents(Precursors) <- factor(Idents(Precursors), levels= my_levels)
All_pre.markers <- FindAllMarkers(Precursors, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

# Select top10 genes
```{r}
top10 <- All_pre.markers %>% group_by(cluster) %>% top_n(10, avg_logFC)
utop10<-unique(top10$gene)
color<-brewer.pal(n = 9, name = "RdBu")
```


# Plot top10 signature genes of each cluster
```{r}
DotPlot(Precursors, features=utop10,  dot.scale = 8) +
  scale_color_gradientn(colours  = rev(color)) + coord_flip() + scale_y_discrete(position = "right") +
  theme( axis.text.x = element_text(angle = 60, hjust = 0))
```



## Cell proportion
#Load cell proportion table
```{r}
table<-read.csv2('Prop_Celltypes_github.csv')
```


## Relative cell porportions
# Relative monocyte subtypes proportions (percentage)
```{r}
table$rel_CMPs<-(table$CMPs.1/table$Total.cells)*100
table$rel_GMPs<-(table$GMPs.1/table$Total.cells)*100
table$rel_HSCs<-(table$HSCs.1/table$Total.cells)*100
table$rel_MEPs<-(table$MEPs.1/table$Total.cells)*100
```


# Plot monocyte subtypes proportions per pseudotime
```{r}
table$Pseudotime<-as.character(table$Pseudotime)
p1 <- ggplot(table, aes(x=Pseudotime, y=rel_CMPs,  group=Pseudotime, color=Pseudotime)) 
p1 <- p1 +  geom_boxplot()
p1 <- p1 + geom_point(aes(color = Pseudotime, group=Pseudotime),na.rm = TRUE, size=8)
p1<- p1 + theme_bw() + theme(axis.text.y = element_text(hjust = 1, size=12, color = "black"), 
                             axis.text.x = element_text(size=14, color = "black"), 
                             axis.title = element_text(size = 20))
p1<- p1 + scale_x_discrete(limits=c("0","1",'2',"3",
                                    "4","5","6",'7'))
p1 <- p1 + scale_color_manual(values = c("#A7A9AC",'#E84F8C',"#7031B9","#A04E9E", "#E65826", "#F99B1C", "#FDC077", "#51BBFE"))
p1 <- p1 + theme_bw() +theme(legend.position = "none",
                             text = element_text(size=16),
                             axis.title.x=element_blank(),
                             axis.text.x=element_blank(),
                             axis.ticks.x=element_blank())
p1 <- p1 + ggtitle("CMPs")

p2 <- ggplot(table, aes(x=Pseudotime, y=rel_GMPs,  group=Pseudotime, color=Pseudotime)) 
p2 <- p2 +  geom_boxplot()
p2 <- p2 + geom_point(aes(color = Pseudotime, group=Pseudotime),na.rm = TRUE, size=8)
p2 <- p2 + theme_bw() + theme(axis.text.y = element_text(hjust = 1, size=12, color = "black"), 
                              axis.text.x = element_text(size=14, color = "black"), 
                              axis.title = element_text(size = 20))
p2<- p2 + scale_x_discrete(limits=c("0","1",'2',"3",
                                    "4","5","6",'7'))
p2 <- p2 + scale_color_manual(values = c("#A7A9AC",'#E84F8C',"#7031B9","#A04E9E", "#E65826", "#F99B1C", "#FDC077", "#51BBFE"))
p2 <- p2 + theme_bw() +theme(legend.position = "none",
                             text = element_text(size=16),
                             axis.title.x=element_blank(),
                             axis.text.x=element_blank(),
                             axis.ticks.x=element_blank())
p2 <- p2 + ggtitle("GMPs")

p3 <- ggplot(table, aes(x=Pseudotime, y=rel_HSCs,  group=Pseudotime, color=Pseudotime)) 
p3 <- p3 +  geom_boxplot()
p3 <- p3 + geom_point(aes(color = Pseudotime, group=Pseudotime),na.rm = TRUE, size=8)
p3 <- p3 + theme_bw() + theme(axis.text.y = element_text(hjust = 1, size=12, color = "black"), 
                              axis.text.x = element_text(size=14, color = "black"), 
                              axis.title = element_text(size = 20))
p3<- p3 + scale_x_discrete(limits=c("0","1",'2',"3",
                                    "4","5","6",'7'))
p3 <- p3 + scale_color_manual(values = c("#A7A9AC",'#E84F8C',"#7031B9","#A04E9E", "#E65826", "#F99B1C", "#FDC077", "#51BBFE"))
p3 <- p3 + theme_bw() +theme(legend.position = "none",
                             text = element_text(size=16),
                             axis.title.x=element_blank(),
                             axis.text.x=element_blank(),
                             axis.ticks.x=element_blank())
p3 <- p3 + ggtitle("HSCs")

p4 <- ggplot(table, aes(x=Pseudotime, y=MEPs.1,  group=Pseudotime, color=Pseudotime)) 
p4 <- p4 +  geom_boxplot()
p4 <- p4 + geom_point(aes(color = Pseudotime, group=Pseudotime),na.rm = TRUE, size=8)
p4 <- p4 + theme_bw() + theme(axis.text.y = element_text(hjust = 1, size=12, color = "black"), 
                              axis.text.x = element_text(size=14, color = "black"), 
                              axis.title = element_text(size = 20))
p4<- p4 + scale_x_discrete(limits=c("0","1",'2',"3",
                                    "4","5","6",'7'))
p4 <- p4 + scale_color_manual(values = c("#A7A9AC",'#E84F8C',"#7031B9","#A04E9E", "#E65826", "#F99B1C", "#FDC077", "#51BBFE"))
p4 <- p4 + theme_bw() +theme(legend.position = "none",
                             text = element_text(size=16),
                             axis.title.x=element_blank(),
                             axis.text.x=element_blank(),
                             axis.ticks.x=element_blank())
p4 <- p4 + ggtitle("MEPs")


CombinePlots(plots = list(p1,p2,p3,p4), ncol=1)
```



## Different cell proportions through pseudotimes
# select patients with covid19
```{r}
table2<-subset(table, table$Pseudotime %in% c("1",'2',"3",
                                              "4","5","6",'7'))
```

# Longitudinal liner mixed model
```{r}
fit <- lmer(rel_HSCs ~ Pseudotime + (1|Patient), data=table2, REML = FALSE)
reduced.fit<- lmer(HSCs.1 ~ 1 + (1|Patient),  data=table2, REML = FALSE)
anova(reduced.fit,fit )
```


# select control patients
```{r}
table_healthy<-subset(table, table$Pseudotime %in% c('0'))
```

#Mann-whitney test comparing healthy with covid19 patients
```{r}
wilcox.test(table2$rel_HSCs, table_healthy$rel_HSCs)
```


# Genes of interest
```{r}
Idents(Precursors)<-Precursors$Pseudotime
my_levels <- c("0","1",'2',"3",
               "4","5","6",'7')
Idents(Precursors) <- factor(Idents(Precursors), levels= my_levels)

VlnPlot(Precursors, features =c('BST2','IFITM1', 'IFITM2', 'IFI6'),  pt.size = 0,ncol = 2,
        cols = c("#A7A9AC", '#E84F8C', "#7031B9", "#A04E9E",'#E65826',"#F99B1C",'#FDC077','#51BBFE'))
```


