---
title: "Megakaryocytes_analysis"
author: "Joana P. Bernardes"
date: "11/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(devtools)
library(SingleR)
library(knitr)
library(RColorBrewer)
library(org.Hs.eg.db)
library(topGO)
library(plyr)
```

##Load object extracted from merged object
```{r}
Megakaryocytes_pre <- readRDS('Megakaryocytes.rds')
```


## Seurat analysis
# Normalize object
```{r}
Megakaryocytes_pre <- NormalizeData(Megakaryocytes_pre)
```

# Scale object using all genes
```{r}
all.genes<-rownames(Megakaryocytes_pre)
Megakaryocytes_pre <- ScaleData(Megakaryocytes_pre, features = all.genes)
```

## Find object variable genes 
```{r}
Megakaryocytes_pre<-FindVariableFeatures(Megakaryocytes_pre, selection.method = "vst", nfeatures = 2000)
```

# Run PCA
# Use elbow plot to define number of dimensions
```{r}
Megakaryocytes_pre <- RunPCA(Megakaryocytes_pre,npcs = 80, ndims.print = 1:5)
ElbowPlot(Megakaryocytes_pre, ndims = 80)
```

# run UMAP
```{r}
Megakaryocytes_pre <- RunUMAP(object = Megakaryocytes_pre, dims = 1:60)

```



# plot basic UMAPs
```{r}
p1<-DimPlot(object = Megakaryocytes_pre, reduction = "umap", pt.size = 2,label = TRUE, group.by = "orig.ident")

p2<-DimPlot(object = Megakaryocytes_pre, reduction = "umap", pt.size = 2,label = TRUE, group.by = "Patient") 

color<-brewer.pal(n = 11, name = "BrBG")
p3<-DimPlot(object = Megakaryocytes_pre, reduction = 'umap', pt.size = 2, label = FALSE, group.by ="Timepoint",
            order = c('TA', 'TA2', 'TB', 'TC', 'TE', 'rec'),
            cols=rev(color))

p4<-DimPlot(object = Megakaryocytes_pre, reduction = "umap", pt.size = 2, label = F, group.by = "Pseudotime",
            cols = c('#A7A9AC',"#E84F8C","#7031B9", '#A04E9E',"#E65826","#F99B1C","#FDC077",  '#51BBFE'))

p5<-DimPlot(object = Megakaryocytes_pre, reduction = "umap" ,pt.size = 2,label = TRUE, group.by = "dmap.labels",
            cols = singleR_colors) 
CombinePlots(plots = list(p1,p2,p3,p4, p5), ncol=1)
```


## Clustering
# Find neighbours
```{r}
Megakaryocytes_pre <- FindNeighbors(Megakaryocytes_pre,  dims = 1:60)
```

# Calculate clusters
```{r}
Megakaryocytes_pre <- FindClusters(Megakaryocytes_pre, resolution = 0.1)
```

# select Megakaryocytes
```{r}
Idents(Megakaryocytes_pre)<-Megakaryocytes_pre$dmap.labels
Megakaryocytes<-subset(Megakaryocytes_pre, idents="Megakaryocytes")
```


# Difusion UMAPs based on Pseudotime
```{r}
p0<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap", pt.size =1,
       order=c( "0",'7',"1",  "3", "4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey',"grey",'grey','black'))
p1<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "1","0",'7', "3", "4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey','grey','grey',"#E84F8C"))
p2<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( '2',"0",'7',"1", "3", "4","5","6" ),
     cols = c( "grey", "grey", "grey",'grey',"grey",'grey','grey', "#7031B9"))
p3<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "3", "0",'7',"1","4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey","grey",'grey','grey', '#A04E9E'))
p4<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "4","0",'7',"1", "3", "5","6", '2'),
     cols = c("grey", "grey", "grey", 'grey',"grey",'grey','grey', "#E65826"))
p5<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "5","0",'7',"1","3","4","6", '2'),
     cols = c( "grey", "grey", "grey",'grey',"grey",'grey','grey', "#F99B1C"))
p6<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "6","0",'7',"1", "3",   "4","5", '2'),
     cols = c("grey",  "grey", "grey",'grey',"grey",'grey','grey', "#FDC077"))
p7<-DimPlot(Megakaryocytes, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c('7', "0","1","3","4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey',"grey",'grey', '#51BBFE'))
CombinePlots(plots = list(p0,p1,p2,p3,p4, p5, p6,p7), ncol=2)
```


# Identify signature genes per Pseudotime
```{r}
Idents(Megakaryocytes)<-Megakaryocytes$Pseudotime
my_levels <- c("0","1",'2', "3", "4","5","6", '7')
Idents(Megakaryocytes) <- factor(Idents(Megakaryocytes), levels= my_levels)
All_pre.markers <- FindAllMarkers(Megakaryocytes, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

#Remove IG genes and select top10 genes
```{r}
All_pre.markers_2<-subset(All_pre.markers, !grepl(("^IG"), All_pre.markers$gene))

top10 <- All_pre.markers_2 %>% group_by(cluster) %>% top_n(10, avg_logFC)
utop10<-unique(top10$gene)
color<-brewer.pal(n = 9, name = "RdBu")
```


# Plot top10 signature genes of each cluster
```{r}
DotPlot(Megakaryocytes, features=utop10,  dot.scale = 8) +
  scale_color_gradientn(colours  = rev(color)) + coord_flip() + scale_y_discrete(position = "right") +
  theme( axis.text.x = element_text(angle = 60, hjust = 0))
```


## Differantial expressed genes during covid-19
```{r}
Idents(Megakaryocytes)<-Megakaryocytes$Pseudotime
Cluster.response1 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "1", verbose = FALSE,test.use = 'MAST' )
Cluster.response1<-subset(Cluster.response1, Cluster.response1$p_val_adj < 0.05)
Cluster.response1$Gene<-rownames(Cluster.response1)
Cluster.response1$Gene=NULL
Cluster.response_Down<-subset(Cluster.response1, Cluster.response1$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response1, Cluster.response1$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 1'
Cluster.response_Down$signal<-'Upregulated in 1'
Cluster.response1<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response2 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "2", verbose = FALSE,test.use = 'MAST' )
Cluster.response2<-subset(Cluster.response2, Cluster.response2$p_val_adj < 0.05)
Cluster.response2$Gene<-rownames(Cluster.response2)
Cluster.response2$Gene=NULL
Cluster.response_Down<-subset(Cluster.response2, Cluster.response2$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response2, Cluster.response2$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 2'
Cluster.response_Down$signal<-'Upregulated in 2'
Cluster.response2<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response3 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "3", verbose = FALSE,test.use = 'MAST' )
Cluster.response3<-subset(Cluster.response3, Cluster.response3$p_val_adj < 0.05)
Cluster.response3$Gene<-rownames(Cluster.response3)
Cluster.response3$Gene=NULL
Cluster.response_Down<-subset(Cluster.response3, Cluster.response3$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response3, Cluster.response3$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 3'
Cluster.response_Down$signal<-'Upregulated in 3'
Cluster.response3<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response4 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "4", verbose = FALSE,test.use = 'MAST' )
Cluster.response4<-subset(Cluster.response4, Cluster.response4$p_val_adj < 0.05)
Cluster.response4$Gene<-rownames(Cluster.response4)
Cluster.response4$Gene=NULL
Cluster.response_Down<-subset(Cluster.response4, Cluster.response4$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response4, Cluster.response4$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 4'
Cluster.response_Down$signal<-'Upregulated in 4'
Cluster.response4<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response5 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "5", verbose = FALSE,test.use = 'MAST' )
Cluster.response5<-subset(Cluster.response5, Cluster.response5$p_val_adj < 0.05)
Cluster.response5$Gene<-rownames(Cluster.response5)
Cluster.response5$Gene=NULL
Cluster.response_Down<-subset(Cluster.response5, Cluster.response5$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response5, Cluster.response5$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 5'
Cluster.response_Down$signal<-'Upregulated in 5'
Cluster.response5<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response6 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "6", verbose = FALSE,test.use = 'MAST' )
Cluster.response6<-subset(Cluster.response6, Cluster.response6$p_val_adj < 0.05)
Cluster.response6$Gene<-rownames(Cluster.response6)
Cluster.response6$Gene=NULL
Cluster.response_Down<-subset(Cluster.response6, Cluster.response6$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response6, Cluster.response6$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 6'
Cluster.response_Down$signal<-'Upregulated in 6'
Cluster.response6<-rbind(Cluster.response_Up, Cluster.response_Down)

Cluster.response7 <-FindMarkers(Megakaryocytes, ident.1 = "0", ident.2 = "7", verbose = FALSE,test.use = 'MAST' )
Cluster.response7<-subset(Cluster.response7, Cluster.response7$p_val_adj < 0.05)
Cluster.response7$Gene<-rownames(Cluster.response7)
Cluster.response7$Gene=NULL
Cluster.response_Down<-subset(Cluster.response7, Cluster.response7$avg_logFC < 0)
Cluster.response_Up<-subset(Cluster.response7, Cluster.response7$avg_logFC > 0)
Cluster.response_Up$signal<-'Downregulated in 7'
Cluster.response_Down$signal<-'Upregulated in 7'
Cluster.response7<-rbind(Cluster.response_Up, Cluster.response_Down)

# Merge Differentially expressed genes
Dif_Exp<-rbind(Cluster.response1, Cluster.response2, Cluster.response3, Cluster.response4, Cluster.response5, Cluster.response6, Cluster.response7)
```


## GO enrichment analysis (example)
#Subset genes
```{r}
table<-subset(Dif_Exp, Dif_Exp$signal %in% 'Upregulated in 1')
```

#TopGO analysis
```{r}
Genes=rownames(table)
anno <- AnnotationDbi::select(org.Hs.eg.db, 
                              keys=rownames(Megakaryocytes), 
                              columns=c("SYMBOL","SYMBOL", "GENENAME", "ENSEMBL"),
                              keytype="SYMBOL")
anSig1 <- as.data.frame(subset(anno, SYMBOL %in% Genes))

genes<-rownames(Megakaryocytes)
overallBaseMean=as.matrix(genes)
rownames(overallBaseMean)<-overallBaseMean
onts = c( "BP")
geneIDs = rownames(overallBaseMean)
inUniverse = geneIDs %in% c(rownames(overallBaseMean)) 
inSelection =  geneIDs %in% anSig1$SYMBOL 
alg <- factor( as.integer( inSelection[inUniverse] ) )
names(alg) <- geneIDs[inUniverse]
tab = as.list(onts)
names(tab) = onts

for(i in 1:3){
  
  ## prepare data
  tgd <- new( "topGOdata", ontology=onts[i], allGenes = alg, nodeSize=5,
              annot=annFUN.org, mapping="org.Hs.eg.db", ID = "symbol" )
  
  ## run tests
  resultTopGO.elim <- runTest(tgd, algorithm = "elim", statistic = "Fisher" )
  resultTopGO.classic <- runTest(tgd, algorithm = "classic", statistic = "Fisher" )
  
  ## look at results
  tab[[i]] <- GenTable( tgd, Fisher.elim = resultTopGO.elim, 
                        Fisher.classic = resultTopGO.classic,
                        orderBy = "Fisher.classic" , topNodes = 200)
  
}

```

# Write results in table
```{r}
topGOResults <- rbind.fill(tab)

for(i in 1:length(topGOResults$GO.ID)){
  go_id <- as.vector(topGOResults[i,1])
  alleges <- get(go_id, org.Hs.egGO2ALLEGS)
  genes <- unlist(mget(alleges, org.Hs.egSYMBOL))
  #print(as.vector(genes))
  genes_in_cat <- intersect(as.vector(genes), as.vector(Genes))
  #print(genes_in_cat)
  gene_sym_in_cat <- as.vector(unlist(mget(unlist(mget(genes_in_cat, 
                                                       org.Hs.egSYMBOL2EG)), org.Hs.egSYMBOL)))
  gene_sym_in_cat_str <- ""
  
  if(length(genes_in_cat) > 0){
    for(j in 1:length(gene_sym_in_cat)){
      gene_sym_in_cat_str <- paste(gene_sym_in_cat_str, 
                                   gene_sym_in_cat[j], sep = ',')
    }
  }
  
  topGOResults$Genes[i] <- gene_sym_in_cat_str
  topGOResults$no_of_genes[i] <- length(gene_sym_in_cat)
}
```


# select top 10 GO terms
```{r}
topGOResults_top <- topGOResults[order(topGOResults$Fisher.classic),] 
topGOResults_top<-head(topGOResults_top,10)

# GOs of interest
plot_data<-read.csv2('GO_interest_Megakaryocytes.csv')
plot_data$Gene.ratio<-plot_data$Significant/plot_data$Annotated
plot_data$Pvalue<--1*log10(plot_data$Fisher.classic)
```

# Plot top GO terms
```{r}
p <- ggplot(plot_data, mapping = aes(x=group, y=reorder(Term, Order), size=Gene.ratio, color=Pvalue))
p <- p + geom_point()
p <- p + scale_x_discrete(limits=c('1', '2', '3', '4' , '5', '6'))
p <- p + xlab("Pseudotime") + ylab("Term")
p <- p + scale_colour_gradient(high="#990000", low="#FF9999", name='P-value')
p <- p + theme_bw() + theme(axis.text.y = element_text(hjust = 1, size=12, color = "black"), 
                            axis.text.x = element_text(size=14, color = "black", angle=45, hjust=1), 
                            axis.title = element_text(size = 20))
p
```


## Genes of interest
# Cohort 1
```{r}
VlnPlot(Megakaryocytes, features =c('ODC1','TGFB1',"IFITM3", 'CD27', 'PKM', 'FCER1G','TREML1', 'ITGA2B', 'IFITM2', 'IFI27', 'TNFSF13'),  pt.size = 0,ncol = 6,
        cols = c("#A7A9AC", '#E84F8C', "#7031B9", "#A04E9E",'#E65826',"#F99B1C",'#FDC077','#51BBFE'))
```

# Difusion UMAPs based on Genes of interest
```{r}
p0<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap", pt.size =1,
       order=c( "0",'7',"1",  "3", "4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey',"grey",'grey','black'))
p1<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "1","0",'7', "3", "4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey','grey','grey',"#E84F8C"))
p2<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( '2',"0",'7',"1", "3", "4","5","6" ),
     cols = c( "grey", "grey", "grey",'grey',"grey",'grey','grey', "#7031B9"))
p3<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "3", "0",'7',"1","4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey","grey",'grey','grey', '#A04E9E'))
p4<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "4","0",'7',"1", "3", "5","6", '2'),
     cols = c("grey", "grey", "grey", 'grey',"grey",'grey','grey', "#E65826"))
p5<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "5","0",'7',"1","3","4","6", '2'),
     cols = c( "grey", "grey", "grey",'grey',"grey",'grey','grey', "#F99B1C"))
p6<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c( "6","0",'7',"1", "3",   "4","5", '2'),
     cols = c("grey",  "grey", "grey",'grey',"grey",'grey','grey', "#FDC077"))
p7<-DimPlot(Plasma, group.by  = 'Pseudotime', reduction = "umap",pt.size =1,
       order=c('7', "0","1","3","4","5","6", '2'),
     cols = c("grey", "grey", "grey", "grey",'grey',"grey",'grey', '#51BBFE'))
CombinePlots(plots = list(p0,p1,p2,p3,p4, p5, p6,p7), ncol=2)
```

